---
title: "Global Trends of CO2 Emission and Effects on the Global Hapinnes Index"
subtitle: "Data Analysis Project - Digital Sciences for Hightech"
author:
- "Guy Krothammer, Yarden Katash, Guy Dahan"
output: 
  html_notebook:
    toc: true
    number_section: false
    fig_width: 10
    fig_height: 10
    fig_caption: true
    df_print: kable
    fontsize: 11pt
    theme: united
---

# Introduction

## Background

This project was done as a part of "Data Analysis" course as part of our teams 
studies in Digital Sciences for High-Tech in the University of Tel-Aviv. 
Our team has great interest in using our studies for exploring and in the future maybe even
developing tools for improving the way we treat our planet. Therefore our subject is
CO2 emissions in congestion with the world happinnes index. 

## Goals

1. Find the trends of CO2 emissions in different countries along the years 1700-2017.
2. Find what countries had greater growth in emissions and in which years.
3. Compare the leading countries with the highest recent trends of growth or decreased emission
in correlation with the current and previous UN happines Index. 

# Part 1: Importing Data and Packages

## Packages Import

```{r}

# install and import packages
install.packages("ggchicklet", repos = "https://cinc.rud.is") # for rounded corners 
install.packages("cowplot") #patch plots
install.packages("hrbrthemes", repos = c("https://cinc.rud.is", "https://cloud.r-project.org/")) #beautiful theme
install.packages("waffle", repos = "https://cinc.rud.is") #waffle plot
install.packages("ggalt") #dumbell plot
install.packages("GGally") # correlation matrix
install.packages("remotes") # to install the relative weight package
remotes::install_github("martinctc/rwa") # relative weight analysis

# import library
library(tidyverse) 
library(ggchicklet)
library(cowplot)
library(hrbrthemes)
library(ggalt)
library(GGally)
library(rwa)
```

## Data import

```{r}
happy_index_2021 <- "https://www.kaggle.com/ajaypalsinghlo/world-happiness-report-2021/download"
emission <- "https://www.kaggle.com/yoannboyere/co2-ghg-emissionsdata/download"

# df_2005 <- read.csv("../Data/world-happiness-report-2005-2020.csv")
df_2021 <- read.csv(happy_index_2021)
df_emiss <- read.csv(emission)
head(df_emiss)
```
# Part 2: Tidying the Data

## 

```{r}

```

# Part 3: Proccesing of the Data

## Visualisation 

### 3.1 
```{r}


```

### 3.2 

```{r}


```

### 3.3

```{r}

```

## Modelling
