---
title: "Global Trends of CO2 Emission and Effects on the Global Hapinnes Index"
author: "Guy Krothammer, Yarden Katash, Guy Dahan"
subtitle: Data Analysis Project - Digital Sciences for Hightech
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    number_section: no
    df_print: kable
    fontsize: 11pt
    theme: united
    code_folding: hide
editor_options:
  chunk_output_type: inline
---

# Introduction

## Background

This project was done as a part of "Data Analysis" course as part of our teams 
studies in Digital Sciences for High-Tech in the University of Tel-Aviv. 
Our team has great interest in using our studies for exploring and in the future maybe even
developing tools for improving the way we treat our planet. Therefore our subject is
CO2 emissions in congestion with the world happiness index. 

## Goals

1. Find the trends of CO2 emissions in different countries along the years 1700-2017.
2. Find what countries had greater growth in emissions and in which years.
3. Compare the leading countries with the highest recent trends of growth or decreased emission
in correlation with the current and previous UN happiness Index. 

# Part 1: Importing Data and Packages

```{r, include=FALSE}
knitr::opts_chunk$set(fig.width=600, fig.height=400, fig.align = "center", warning = FALSE, message = FALSE) 
```

```{r, include=FALSE}
# install and import packages

install.packages(c("cowplot", "ggalt", "ggalt", "GGally", "Hmisc", "ISLR", "skimr", "ggcorrplot", "ggpubr", "DiagrammeR", "xts", "countrycode", "countrycode", "shiny"))

remotes::install_github("hrbrmstr/streamgraph")
remotes::install_github("martinctc/rwa") # r
install.packages("waffle", repos = "https://cinc.rud.is") #waffle plot
install.packages("ggchicklet", repos = "https://cinc.rud.is") # for rounded corners 
```

```{r, include=FALSE}
# import library
library(tidyverse)
library(ggchicklet)
library(cowplot)
library(hrbrthemes)
library(ggalt)
library(GGally)
library(rwa)
library(readr)
library(Hmisc)
library(tibble)
library(dplyr)
library(ISLR)
library(skimr)
library(ggcorrplot)
library(ggpubr)
library(xts)
library(countrycode)
library(plotly)
library(shiny)
library(streamgraph)
```


## Data import

```{r, results='hide'}
happy_index_2005 <- "Data/world-happiness-report-2005-2020.csv" 
happy_index_2021 <- "Data/world-happiness-report-2021.csv"
emission <- "Data/co2_emission.csv"
population <-"Data/population.csv"

df_2005 <- read.csv(happy_index_2005)
df_2021 <- read.csv(happy_index_2021)
df_emiss <- read.csv(emission, col.names = c("Entity", "Code" , "Year", "co2"))
df_pop <- read.csv(population, col.names = c("Entity", "Code" , "Year", "Population"))
```

The first Data consists of the UN happiness index for the years 2005-2017.

```{r, collapse=TRUE}
as_tibble(df_2005)
```
The second data is the happiness index report for 2021.

```{r, collapse=TRUE}
as_tibble(df_2021)
```
The third data contains CO2 emissions from around 1700 up until 2017, by countries and continents.

```{r, collapse=TRUE}
as_tibble(df_emiss)
```
The final data is a list of population sizes by countries for about the same years as the CO2 data.

```{r, collapse=TRUE}
as_tibble(df_pop)
```

As seen in the table above, the main dataset regarding emissions has only 4 features, of which 2 are identical (countries and country code). We will now examine the summary of it's characteristics:

```{r}
summary(df_emiss)
```
As seen above, we can see the minimal year, minimal amount of CO2 emissions. Also shown are mean and max values. 
We also learn about the size of the data, around ~21k rows. 

# Part 2: Tidying the Data

## Tidying Functions

```{r, echo=FALSE}

# add "dif" column

df_emiss <-df_emiss %>% 
  group_by(Entity) %>%
  mutate(Diff = co2 - lag(co2, default = co2[1]))

#filter year up 1950

df_pop_to_merge <- df_pop %>%
  filter(Year > 1949)

df_emiss_to_merge <- df_emiss %>%
  filter(Year > 1949)

#arrange data to merge emission and population 

df_emiss_to_merge$Entity <- as.character(df_emiss_to_merge$Entity)
df_pop_to_merge$Entity <- as.character(df_pop_to_merge$Entity)

df_all <- full_join(df_emiss_to_merge, df_pop_to_merge, by = c('Entity','Year'))

drops <- c("Code.y")
df_all <-df_all[, !(names(df_all) %in% drops)]

names(df_all)[names(df_all) == "Code.x"] <- "Code"

#add column of normalized emission: co2/population by year (avg.ratio)

df_all <- df_all %>%
  mutate(ratio = co2/Population)

### Create dataframe for both 30 years average and by countries

names(df_2005)
names(df_2005)[names(df_2005) == "ï..Country.name"] <- "Entity"

df_2017 <- df_2005 %>%
  filter(year == 2017)%>%
  select(c(Life.Ladder, Entity))

df_last_30 <- df_all %>%
  filter(Year >= 1987) %>%
  na.omit()

agg1 <- aggregate(df_last_30[,c('ratio', 'Population', 'co2')], 
                  list(df_last_30$Entity), mean) %>%
  set_names(nm=c('Entity', 'avg.ratio', 'avg.Population', 'avg.co2'))

agg2 <- aggregate(df_last_30[,c('Diff')],
                  list(df_last_30$Entity), sum ) %>%
  set_names(nm=c('Entity', 'sum.dif'))

df_2017$Entity <- as.character(df_2017$Entity)
agg1$Entity <- as.character(agg1$Entity)
agg2$Entity <- as.character(agg2$Entity)

df_merge <- full_join(agg1, df_2017, by = c('Entity')) %>%
  full_join(agg2) %>%
  na.omit() # TODO check why 20 countries were removed

df_merge
```

We have just created the two main datasets to work with, which includes important features from all of the previous datasets and some features that we have created above:

1. "df_all" which include all data per country per year from 1950 til 2017, introducing some new features:
  * Population, taken from the the population's DF
  * CO2/Population ratio, calculated.
  * "Diff" column, which shows the difference between the previous and the current year emissions for each consecutive year.
  
2. "df_merge", which includes data for each country regarding:
  * The average ratio feature, which is calculated from 1950 to 2017 using the previously explained "ratio" feature. 
  * The difference in co2 emission this period and the life-ladder in 2017.
  * The average population in those years. 
  * The average CO2 emissions.
  * The 2017's "Life.Ladder" - happiness index score. 
  * "Sum.Dif" - summing over the diff feature. 

We also made a choice to limit our data regarding emissions to after 1950. The reason is that the rise in values is almost exponential after the years of WW2. When plotted on a graph, placing values from before and after 1950 they become incomparable.  

The summary for df_merge:

```{r, collapse=TRUE}
summary(df_merge)
```
And for df_all:
```{r, collapse=TRUE}
summary(df_all)
```
# Part 3: Proccesing of the Data

## Visualisation 

### Countries of the World
In this plot the different countries in the data are shown in deeper colors for higher values of emission to population ratio in the year of 2017

```{r, fig.align='center', warning=FALSE, message=FALSE}
d = df_all %>%
  filter(Year==2017)

l <- list(color = toRGB("grey"), width = 0.2)

# specify map projection/options

g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

p <- plot_geo(d) %>%
  add_trace(
    z = ~ratio, color = ~ratio, colors = 'Reds',
    text = ~Entity, locations = ~Code, marker = list(line = l)
  ) %>%
  colorbar(title = 'CO2/Population', thickness=15) %>%
  layout(
    title = 'World Ratio of CO2/Population',
    geo = g,
    autosize = F
  )
```


```{r, fig.align='center', warning=FALSE, message=FALSE}
p <- ggplotly(p, width = 3000, height = 1500, automargin = TRUE)
p
```
As we can see, some of the leading countries in CO2/Population ratio are Qatar, United Arab Emirates, Kuwait, USA, Saudi Arabia, Canada, Kazakhstan and Australia with above 15 units of CO2/population.
That means countries like China and India - whose CO2 emission rate is very high compared to other countries - are ranked lower, even though they are more polluting in net amount.

### Countries Emission vs Population by Years and continents 

For comfortably displaying the data, each country is allocated to it's fitting continent and a color. 

```{r, include=FALSE, message=FALSE, warning=FALSE}
d <- df_all %>%
  filter(Year > 2000) %>%
  filter(Entity != 'World')

d_cont<- data.frame(country = c(d$Entity))

d_cont$continent <- countrycode(sourcevar = d_cont[, 'country'],
                            origin = "country.name",
                            destination = "continent")

d$continent <- d_cont$continent

d <- na.omit(d)
```


```{r, message=FALSE, collapse=TRUE, warning=FALSE}
fig <- d %>%
  plot_ly(
    x = ~co2, 
    y = ~Population, 
    size = ~ratio,
    frame = ~Year, 
    text = ~Entity, 
    color = ~continent,
    type = 'scatter',
    mode = 'markers',
    height = 500,
    width = 900,
    automargin = TRUE 
  )

fig <- fig %>% layout(
      title = "Yearly CO2 Emissions by Countries vs Population",
      xaxis = list(
      type = "log"
      # autosize = F
    )
  )

fig
```
We would like to develop this graph forward and see it with correlation to the life ladder index from 2005-2017. For this purpose we would like to create another df, containing both happiness index scores and data from "df_all", which contains continents data as well now: 

```{r, echo=FALSE}
df_2005_2017 <- d %>%
  filter(Year >= 2005) %>%
  na.omit()
temp <- df_2005[,c('Entity', 'Life.Ladder', 'year')] 
colnames(temp)[colnames(temp) == "year"] <- "Year"

df_2005_2017 <-df_2005_2017%>%
  full_join(temp, by=c("Year", "Entity"))
df_2005_2017

df_2005_2017$Population <- as.numeric(df_2005_2017$Population)
```
We get the following plot:
```{r, collapse=TRUE, warning=FALSE, message=FALSE, echo=TRUE}
fig <- plot_ly(
    data = df_2005_2017,
    x = ~co2, 
    y = ~Life.Ladder, 
    size = ~ratio,
    frame = ~Year, 
    text = ~Entity, 
    color = ~continent,
    type = 'scatter',
    mode = 'markers',
    height = 500,
    width = 900,
    automargin = TRUE 
  )

fig <- fig %>% layout(
      title = "Yearly CO2 Emissions by Countries vs Life Ladder",
      xaxis = list(
      type = "log"
    )
  )

fig
```
And yet another scatter plot for different countries:
```{r, warning=FALSE, message=FALSE}
df_3d <- df_2005_2017 %>%
  filter(Year == 2017)

fig <- df_3d %>%
  plot_ly(
    x = ~Population, 
    y = ~ratio, 
    z = ~Life.Ladder,
    # size = ~Life.Ladder,
    # frame = ~Entity, 
    text = ~Entity, 
    color = ~continent,
    height = 500,
    width = 900,
    automargin = TRUE 
  )

fig <- fig %>% layout(
      title = "CO2 Emissions vs Life Ladder, Filtered by Countries, 2017"
    )

fig <- fig %>% add_markers()
fig
```
This graph shows the CO2 emissions of each country compared to it's population and Life Ladder score. If we look at the United States for example, it seems that while the population has increased, the level of pollution has decreased. 

### Commulative Emissions by Countries Over the Years 1950-2017

```{r, message=FALSE}

d <- df_all %>%
  filter(Entity != "World") %>%
  filter(co2 != 0)
d
pp <- streamgraph(d, key="Entity",
                  order = "asis",
                  value="co2", 
                  date="Year",
                  offset="zero",
                  sort="co2"
                  ) %>%
  sg_axis_y(tick_format = "e")  %>%
  sg_legend(show=TRUE, label="Country: ") %>%
  sg_title("CO2 Emissions by Years and Countries, 1950-2017")

pp


```
This plot shows the accumulation of emissions year after year, from 1950 to 2017 for different countries.

### Commulative Happinness Trends in Single Country 

This graph shows the accumulative happiness index for each country from 2005 to the present. 

```{r, message=FALSE, warning=FALSE}
pp <- streamgraph(df_2005, key = "Entity",
                  # order = "reverse",
                  value="Life.Ladder", 
                  date="year"
                  # offset="zero",
                  ) %>%
  sg_fill_brewer("Blues") %>%
  sg_legend(show=TRUE, label="Country: ") %>%
  sg_title("Happiness Index by Years and Countries, 1950-2017") 
  
pp


```

### Yearly in CO2 Emissions for Top 10 Polluting Countries

```{r}
top_emiss_2017 <- df_emiss %>%
  filter(Year == 2017) %>%
  filter(Code != '') %>%
  filter(Entity != "World") %>%
  top_n( 10, co2) %>%
  arrange(desc(co2)) %>%
  head(10)
  
top_countries = top_emiss_2017$Entity
```

```{r}
g <- ggplot(data = df_emiss[df_emiss$Entity %in% top_countries, ] 
                          , aes(x = Year, y = co2, group = Entity)) + 
    geom_line() +
    labs(title = "Top 10 Countries by CO2 Emissions") + 
    geom_line(aes(col = Entity)) +
    theme_ipsum()
g <- ggplotly(g, height = 500,
    width = 900,
    automargin = TRUE)
g
```


```{r}
g <- ggplot(data = df_all[df_all$Entity %in% top_countries, ] 
                          , aes(x = Year, y = ratio, group = Entity)) + 
    geom_line(aes(col = Entity)) + 
    labs(title = "Top 10 Countries by CO2 Emissions to Population ratio") +
    theme_ipsum()

g <- ggplotly(g, height = 500,
    width = 900,
    automargin = TRUE)
g

```
According to these two graphs, using the CO2/Population ratio. For example, Saudi Arabia has a lower population than china, but in terms of pollution per person in 2017 Saudi Arabia had the highest emission per person ratio in the world.

### Boxplot for top 10 Countries 

```{r, collapse=TRUE, message=FALSE, echo=TRUE, warning=FALSE}

d <- df_all[df_all$Entity %in% top_countries, ] 
fig <- plot_ly(d,
               x = ~Entity,
               y = ~Diff,
               type = "box",
               color = ~Year,
               height = 500,
                width = 900,
                automargin = TRUE)

fig <- fig %>%  layout(
  title = "Difference in Emission for Top 10 Polluting Countries",
  boxmode = "group")
fig
  
```
### Top 10 Happiest countries Emmissions by Years

This graph takes the 10 happiest countries and shows the distribution of CO2 emissions difference level of emission between 1950 and 2017.


```{r, collapse=TRUE}

top_10_happiness <- df_merge %>%
  filter(rank(desc(Life.Ladder))<=10)

g <- ggplot(data = df_emiss[df_emiss$Entity %in% top_10_happiness$Entity, ] 
                          , aes(x = Year, y = co2, group = Entity)) + 
    geom_line() +
    labs(title = "Top 10 happiness Countries by CO2 Emissions") + 
    theme_ipsum() +
    geom_line(aes(col = Entity))
g <- ggplotly(g, height = 500,
    width = 900,
    automargin = TRUE)
g
```
In this section we are going to suggest two types of correlation based on our exploration of the data:

1. There is correlation between the total trend of emissions and the current happiness index.
2. There is correlation between current emission normalized by the population to the current happiness index.

```{r, collapse=TRUE}
g <- ggplot(data = df_merge, aes(x = avg.ratio, y = Life.Ladder)) + 
    geom_point(aes(colour = Entity)) + 
    labs(title = "Average Ratio vs Happiness Index") +
    theme_ipsum() 
g <- ggplotly(g, height = 500,
    width = 900,
    automargin = TRUE)
g
```

In the following plot we have removed china and India because they are sort of "outliers" in a sense that they have displayed much larger grow in emissions compared to other countries, therefore making it very hard to display on the same plot.


```{r, collapse=TRUE}
df <- subset(df_merge, Entity != "China" & Entity != "India" )

g <- ggplot(data = df, aes(x = Life.Ladder , y = sum.dif)) + 
    geom_point(aes(colour = Entity)) + 
    labs(title = "Sum of Emissions Difference vs 2017 Happiness Index (China, India ex.)") +
    theme_ipsum()
g <- ggplotly(g, height = 500,
    width = 900,
    automargin = TRUE)
g
```

 # Part 4: Modeling
 
 We are now going to introduce the linear model:
 
```{r, collapse=TRUE}
linearmod  <- lm(Life.Ladder ~ avg.ratio, data=df_merge)

summary(linearmod)
```


```{r, collapse=TRUE, warning=FALSE, message=FALSE}
confint(linearmod, level=0.95)
p <- plot_ly(
  x = fitted(linearmod),
  y = residuals(linearmod),
  height = 500,
    width = 900,
    automargin = TRUE)

p<- p %>% layout(
  title = "Residual plot for the Linear Model",
  autosize = T,
  yaxis = list(title = 'Residuals'),
  xaxis = list(title = 'Fitted Linear Model')
  
)

p
```

# TODO add here EXPLANATION
The residual plot shows no signs of 


```{r, collapse=TRUE}
g <- ggplot(df_merge, aes(x = avg.ratio, y = Life.Ladder)) +
  geom_point(aes(colour = Entity)) + 
  theme_bw() +
  stat_smooth(method = "lm") + 
  labs(title = "Linear Modelling Happiness Index and Average Ratio")
g <- ggplotly(g, height = 500,
    width = 900,
    automargin = TRUE)
g


```

According to the model, it can be seen that the correlation between the happiness index and the CO2 emission is not high, but it does exist. We will discuss it in the conclusions section.


```{r, collapse = TRUE}

numeric_tidy <- df_merge[-1]

corr_data <- cor(numeric_tidy)

g <- ggcorrplot(corr_data, hc.order = TRUE, type = "lower",
   outline.col = "white") +
  labs(title = "Correlation Matrix") +
  theme_ipsum()

ax <- list(
  title = "",
  showgrid = FALSE)
  
g <- ggplotly(g, height = 500,
    width = 900,
    automargin = TRUE) %>%
  layout(xaxis=ax, yaxis=ax)
g


```
We can see in the matrix above that there are some correlation that are easy to explain, such as a pretty strong one between the sum of differences regarding emissions over the test period, to the average size of population. There is also a weaker but existing correlation between the life ladder scale and the average ratio of CO2/Population size. 
The following section presents another model and it's summary:

```{r, collapse=TRUE, warning=FALSE}
gmod  <- lm(Life.Ladder ~ log(avg.ratio), data=df_merge)

summary(gmod)
```


```{r, collapse=TRUE, warning=FALSE}
#plot x

p <- qplot(x = avg.ratio, Life.Ladder, data=df_merge)
p <- p + geom_smooth(method = "glm", formula = y~log(x), family = binomial(link = 'log')) +
  geom_point(aes(colour=Entity)) +
  theme_ipsum() +
  labs(title = "Linear Regression Model - Average Ratio to Life Ladder Score")
p <- ggplotly(p, height = 500,
    width = 900,
    automargin = TRUE)
p
```
On this model, We can see a weak correlation, but an existing one to the logistic regression. Interestingly enough, The "happiest" countries seem to have a higher ratio of CO2 emissions as well. Also worth mentioning from this plot:
* The US is located far right on this map, meaning it's emitting a lot of CO2 in proportion to it's population, but it is pretty high on the happiness index as well. 
* Kuwait, being the leader of this unfortunate characteristic, also has a pretty decent Life Ladder score. 


## Testing the Top 30 Happiest Countries


To further check existing correlation, we would compare the differences between the 30 top rated countries in the happiness index and their avg.ratio of CO2/Population. 
In this case, we have two unrelated (i.e., independent or unpaired) groups of samples. Therefore, it is possible to use an independent t-test to evaluate how their means differ. 

* $mA$ - Weighted average of the 30 most happiest countries.
* $mB$ - Weighted average of the rest of the countries.

Our research questions:

* Is the mean avg.ratio of top 30 happiness ($mA$) is greater than the mean of other countries ($mB$)? 

* Is the weighted average of the 30 happiness countries (mA) greater than the weighted average of the other countries ($mB$)?

H0: $mA ≥ mB $  - The null hypothesis

Ha: $mA < mB $ (less) - The alternative hypothesis


```{r}
top_30_happiness <- df_merge %>%
  filter(rank(desc(Life.Ladder))<=30)

# Create a data frame
T_data <- df_merge %>%
  select(Entity, avg.ratio,) %>%
  mutate(group = ifelse(Entity %in% top_30_happiness$Entity, "Top 30 Contries", "Other Countries")) 

g <- ggboxplot(T_data, x = "group", y = "avg.ratio", 
          color = "group", palette = c("#00AFBB", "#E7B800"),
        ylab = "avg. ratio", xlab = "Group") +
  theme_ipsum() +
  labs(title = "Average Ratio Compared - Top 30 Countries vs Other Countries")

g <- ggplotly(g, height = 500,
    width = 900,
    automargin = TRUE)
g


```

Although not dramatic, it is obvious that some correlation does exist. We would like to further check:

* Do the two group have the same variances?
* We will use F-test to test for homogeneity in variances. 

```{r}
res.ftest <- var.test(avg.ratio ~ group, data = T_data)
res.ftest

```
The p-value of F-test is p = 0.112 which is greater than the significance level 0.05.
In conclusion, there is no significant difference between the variances of the two data sets. Therefore, we used the T-test and assumed that the variances are equal - according to case 2 of hypothesis testing.

```{r}

# Compute t-test
res <- t.test(avg.ratio ~ group, data = T_data, var.equal = TRUE, alternative = "less")
res

```
The p-value of the test is $3.103e-07$, which is less than the significance level $\alpha = 0.05$. We can conclude that top 30 average avg.ratio is significantly different from the other countries average avg.ratio with a $pvalue = 3.103e-07$.

# Conclussion

Generally, regarding our base assumptions, we were surprised to see that they were actually not only wrong, but almost opposite. while very undeveloped countries, like countries in Africa mostly, have a very low pollution index and life ladder index. On the other hand, some very developed countries have higher values on both indexes. There are also a lot of countries who don't fit those assumptions at all. 

Although, our final tests did show some weak correlation within the top 30 countries as shown.  

we stipulate that these weak correlations are caused partially because of the fact that the happiness index is a *survey* based data. 
it means that it is possible that the citizens of a country are not necessarily aware of the CO2 pollution and it's consequences. There are also other factors that affect the happiness index like welfare , corruption,  health services, wealth etc.

On the other hand:
1. It can be assumed that a country that emits more CO2 is a more modern and technologically advanced country, therefore it is more established, its citizens feel a sense of progress, and that causes the happiness index to rise.
2. In more polluting countries - it is a possible that there are more workplaces, therefore these countries will have a lower unemployment rate - which can in turn increase the level of happiness index within the country.


In conclusion, in our opinion, the people may be aware of their country CO2 emissions, yet it doesn't necessarily show in the happiness index survey.

Another thing to consider is that environmental preservation, pollution reduction and global warming are quite trending in the last few years, but this is a very complex system of dependencies, and the rate of which changes are noticeable is rapidly changing. It is unknown whether we have crossed the non-return point for making the planet to hot to live at due to those emissions, but time will tell. 




